Tree growth, mortality, and climatic stress in west-coast states, USA

### 2025-09-12 Manuscript.Rmd
This push contains a few changes for the manuscript and the working, published v1 of the app.

Growth domain comparisons (`r fig_nums("fig_diffgrow")`, `r t.ref("tbl_summ_grow_mort")`) did not appear to be produce substantially more "significant" negative vs. positive estimates, and all three columns of `r fig_nums("fig_diffgrow")` had similar numbers of "significant" comparisons. Growth may have been susceptible (negatively or positively) to a change in CWD and differed by initial CWD values, and the responses are species-specific. *Thuja plicata* exhibited greater growth in the Above and Below Medium domains than in the Lower or Higher domains.  *Abies amablis* had lower growth in the Above and Below High domains than the others, where initial-visit CWD was the greatest.  *Picea englelmannii* had lower growth in the Low domain than the High domain, indicating that more growth was occurring where trees were more water stressed.

Mortality domain comparisons are similar in certain regards to the growth comparisons: results for the three types of domain comparisons did not appear to be predominantly positive or negative (`r fig_nums("fig_diffmort")`, `r t.ref("tbl_summ_grow_mort")`).  There were fewer mortality estimates with "significant" findings than growth. There were fewer differences for "Above vs. Below Threshold" comparisons (`r t.ref("tbl_summ_grow_mort")`) than for the comparisons Above or Below the threshold, indicating that mortality may not be as strongly associated with the change in CWD as by the starting value.  *Abies concolor* had greater mortality at the Above Low and High categories than the Above Middle.  *Calocedrus decurrens* had a higher mortality rate at the High Above domains than the other Above domains. *Picea engelmannii* was the opposite; mortality was lowest at the Above High domain compared to other domains.  *Psuedotsuga menziesii* exhibited greater mortality at its Below High domain than other Below domains.  The two *Quercus* species did not exhibit strong patterns to their mortality.

>>>>

Growth domain comparisons (`r fig_nums("fig_diffgrow")`, `r t.ref("tbl_summ_grow_mort")`) did not appear to be produce substantially more "significant" negative vs. positive estimates, and all three columns of `r fig_nums("fig_diffgrow")` had similar numbers of "significant" comparisons. Growth may have been susceptible (negatively or positively) to a change in CWD and differed by initial CWD values, and the responses are species-specific. *Abies amabilis*, *Abies grandis*, *Calocedrus decurrens*, *Pinus ponderosa*, *Pseudotsuga menziesii*, *Tsuga mertensiana*, *Acer macrophyllum*, and *Alnus rubra* exhibited greater growth in the Low domains than the Medium and/or High domains.  Except for *Alnus rubra*, these results predominantly occurred in both the Below domain (2) or the Above and Below domains (5).  *Thuja plicata* had higher growth in the Above and Below Medium domains than either High or Low domains. *Abies concolor*, *Larix occidentalis*, *Picea engelmannii*, *Pinus contorta*, *Pinus lambertiana*, *Arbutus menziesii*, and *Quercus kelloggii* had greater growth in the more water stressed High domains than the Medium and/or Low domains. Except for *Larix occidentalis*, these results predominantly occurred in both the Above domain (3) or the Above and Below domains (3). Other species exhibited no 'significant' findings or displayed a mixture of results.

Mortality domain comparisons are similar in certain regards to the growth comparisons: results for the three types of domain comparisons did not appear to be predominantly positive or negative (`r fig_nums("fig_diffmort")`, `r t.ref("tbl_summ_grow_mort")`).  There were fewer mortality estimates with "significant" findings than growth. There were fewer differences for "Above vs. Below Threshold" comparisons (`r t.ref("tbl_summ_grow_mort")`) than for the comparisons Above or Below the threshold, indicating that mortality may not be as strongly associated with the change in CWD as by the starting value.  *Larix occidentalis*, *Picea engelmannii*,	*Pinus contorta*,	*Pinus monticola*, *Thuja plicata*,	and *Arbutus menziesii* had greater mortality in the lower water deficit Low domains than Medium and/or High domains. These results occurred in Above (3) or both Above and Below (3) domains. *Abies lasiocarpa*,	*Calocedrus decurrens*,	*Pinus lambertiana*, *Pinus ponderosa*,	*Pseudotsuga menziesii*,	*Tsuga heterophylla*,	*Tsuga mertensiana*,	*Acer macrophyllum*,	and *Notolithocarpus densiflorus* had more mortality in the High domains than the Medium and/or Low domains.  A(1111) both() B(11111)

*Abies concolor* had greater mortality at the Above Low and High categories than the Above Middle.  *Calocedrus decurrens* had a higher mortality rate at the High Above domains than the other Above domains. *Picea engelmannii* was the opposite; mortality was lowest at the Above High domain compared to other domains.  *Psuedotsuga menziesii* exhibited greater mortality at its Below High domain than other Below domains.  The two *Quercus* species did not exhibit strong patterns to their mortality.
